A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-15 survey.tex
Updated the bot openings to better reflect the races of each bot
completed the openings descriptions

In the  end, Skynet  won the competition  with its solid  Protoss play
(results  are  summarized   in  Table  \ref{tab:aiide2011}).  The  bot
executed one of a small set of strategies randomly at the start of the
match based on the map and the race of the opponent. Skynet would then
amass a  medium to large  sized army and  expand before moving  out to
attack. Good  use of Dragoon range and  kiting micromanagement allowed
it to hold off the early aggression of other bots such as UAlbertaBot,
which came in 2nd.

@@

AIUR also chose  Protoss, with a strategy which  was in between Skynet
and UAlbertaBot in terms of  attack timings.  At that time, AIUR chose
one  mood among  five (leading  to slightly  different  strategies and
tactics) at the  beginning of a game and kept it  until the end. These
five moods were:
\begin{itemize}
\item  Rush:  where  the  bot  tries  early  attacks,  and  have  good
  probabilities to send  the two or three first  Zealots to harass the
  opponent.
\item Aggressive: where  we have less chance to  perform harasses with
  the first  Zealots, and  the first attack  is usually a  bit delayed
  with regard to the Rush mood.
\item Macro: where the AI do not try any early attacks and focus a bit
  more on its economy before attacking.
\item Defense: where the AI  ``turtles'' and wait to have a consequent
  army before running an attack.
\item  Fast expand:  where the  first building  constructed it  a base
  expansion, for a very economical-oriented game.
\end{itemize}
Notice that  build orders are not  fully hard-coded since  they can be
altered by AIUR's Spend Manager.

@@

One of the major conclusions from the results of the StarCraft competitions, is the lack of adaptation of bots.  Some switch between different build-orders, but do not fully adapt their strategy. No bot is capable of observing the opponent and autonomously synthesize a good plan from scratch to counter the opponent strategy. In this section we analyzed this claim quantitatively. We analyzed the replays from the 2011 and 2012 AIIDE competitions (shown in Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} respectively). Specifically, we analyzed the way bots choose their openings depending on which other bot they are playing against. Given that there is no dominant strategy in StarCraft, and it is necessary to see what the opponent is doing in order to determine the best opening, we would expect bots to change their openings depending on which other bot they are playing (since each bot uses a different strategy, or set of strategies). Using clustering, we identified the most common openings in all the replays from the 2011 and 2012 competitions (each one shown with a different color in the figures).

Specifically, we identified the following openings (Most of the following names are specific building or units of StarCraft, for a better comprehension we refer the reader to \footnote{\url{http://wiki.teamliquid.net/starcraft/Category:Strategies}}):
\begin{itemize}
  \item Protoss openings:
  \begin{itemize}
    \item \emph{two\_gates}: Build two \emph{Gateways} and keep training \emph{Zealots}.
    \item \emph{fast\_dt}: Produce \emph{Dark Templars} as soon as possible.
    \item \emph{templar}: Train \emph{High Templars} as fast as possible.
    \item \emph{speedzeal}: Train \emph{Zealots}  and research attack and speed upgrades as soon as possible.
    \item \emph{corsair}: Produce \emph{Corsairs} as soon as possible and then make transition to train \emph{Dark Templars}, \emph{Reavers} or \emph{Dragoons}.
    \item \emph{nony}: Build three \emph{Gateways} and massive training of \emph{Dragoons}.
    \item \emph{reaver\_drop}: Train \emph{Reavers} as soon as possible to be able to do drops.
  \end{itemize}
  \item Terran openings:
  \begin{itemize}
    \item \emph{bio}: Produce a large army of \emph{Marines} and \emph{Medics}.
    \item \emph{rax\_fe}: Take the closest ``natural expansion'' as soon as possible. This provides a big economic boost in the early to mid game.
    \item \emph{two\_facto}: Build two \emph{Factories} and keep producing \emph{Tanks}.
    \item \emph{vultures}: Produce mainly \emph{Vultures} and research mines.
    \item \emph{air}: Produce \emph{Wraiths} as soon as possible for an air attack.
    \item \emph{drop}: Train \emph{Dropships} as soon as possible to be able to do drops. 
  \end{itemize}
  \item Zerg openings:
  \begin{itemize}
    \item \emph{speedlings}: Train \emph{Zerlings} and research speed upgrade as soon as possible.
    \item \emph{fast\_mutas}: Produce mainly \emph{Mutalisks}.
    \item \emph{mutas}: Expand two times for a stronger economy before massive training of \emph{Mutalisks}.
    \item \emph{lurkers}: Train \emph{Lurkers} as soon as possible to take benefit of its cloak ability.
    \item \emph{hydras}: Massive production of \emph{Hydralisks}.
  \end{itemize}
\end{itemize}

@@

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botopenings.pdf}
    \caption{Distribution of different openings performed by the different bots participating in the AIIDE 2011 competition. For each bot match-up, the colors show the proportion of times that the column bot used a particular opening against the row bot.}
    \label{fig:aiide2011-botopenings}
\end{figure*}

@@

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botopenings2012.pdf}
    \caption{Distribution of different openings performed by the different bots participating in the AIIDE 2012 competition. For each bot match-up, the colors show the proportion of times that the column bot used a particular opening against the row bot.}
    \label{fig:aiide2012-botopenings}
\end{figure*}

@@

\item {\bf Integration:}
\begin{itemize}
\item Multi-scale planning/reasoning: as described in this paper, all the bots developed for the StarCraft AI competitions decompose the problem of playing an RTS game into smaller subproblems, and then solutions for each of those subproblems are integrated in to a common architecture to play the game. However, the integration of each of the modules in a unified architecture is still an open problem. For example, how can decisions made at high-level modules be integrated with decisions made at lower-level modules? 
\end{itemize}

\item {\bf Domain Knowledge: }
\begin{itemize}
\item We know how to incorporate some aspects of domain knowledge (e.g. build orders) into RTS game playing agents. But, in general, how to incorporate some forms of domain knowledge into algorithms for RTS games is still an open problem. For example, standard techniques to encode strategies for other forms of games, like Behavior Trees, are hard to deploy in RTS games. Is it possible to devise techniques that can automatically mine the existing collections of domain knowledge for an RTS game like StarCraft, and incorporate it into the bot? An initial exploration of this idea was carried out by Branavan et al. \cite{branavan2011learning}.
\end{itemize}
\end{itemize}

>>>>

In the  end, Skynet  won the competition  with its solid  Protoss play
(results  are  summarized   in  Table  \ref{tab:aiide2011}).  The  bot
executed one of a small set of strategies randomly at the start of the
match based on the map and the race of the opponent. Skynet would then
amass a  medium to large  sized army and  expand before moving  out to
attack. Good  use of Dragoon (powerful ranged ground unit with clumsy movement) range and  kiting micromanagement allowed
it to hold off the early aggression of other bots such as UAlbertaBot,
which came in 2nd.

@@

AIUR also chose  Protoss, with a strategy which  was in between Skynet
and UAlbertaBot in terms of  attack timings.  At that time, AIUR chose
one  mood among  five (leading  to slightly  different  strategies and
tactics) at the  beginning of a game and kept it  until the end. These
five moods were:
\begin{itemize}
\item  Rush:  where  the  bot  tries  early  attacks,  and  have  good
  probabilities to send the two or three first Zealots (basic contact attack ground unit) to harass the
  opponent.
\item Aggressive: where  we have less chance to  perform harasses with
  the first Zealots, and the first attack is usually a bit delayed
  with regard to the Rush mood.
\item Macro: where the AI do not try any early attacks and focus a bit
  more on its economy before attacking.
\item Defense: where the AI  ``turtles'' and wait to have a consequent
  army before running an attack.
\item  Fast expand:  where the  first building  constructed it  a base
  expansion, for a very economical-oriented game.
\end{itemize}
Notice that  build orders are not  fully hard-coded since  they can be
altered by AIUR's Spend Manager.

@@

One of the major conclusions from the results of the StarCraft competitions, is the lack of adaptation of bots.  Some switch between different build-orders, but do not fully adapt their strategy. No bot is capable of observing the opponent and autonomously synthesize a good plan from scratch to counter the opponent strategy. In this section we analyzed this claim quantitatively. We analyzed the replays from the 2011 and 2012 AIIDE competitions (shown in Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} respectively). Specifically, we analyzed the way bots choose their openings depending on which other bot they are playing against. Given that there is no dominant strategy in StarCraft, and it is necessary to see what the opponent is doing in order to determine the best opening, we would expect bots to change their openings depending on which other bot they are playing (since each bot uses a different strategy, or set of strategies). Using clustering, we identified the most common openings in all the replays from the 2011 and 2012 competitions (each one shown with a different color in the figures). No data is shown for the ItayUnvermind bot, since its opening did not match significantly with any of the ones used in our study.

Specifically, we identified the following openings (Most of the following names are specific building or units of StarCraft, for a better comprehension we refer the reader to \footnote{\url{http://wiki.teamliquid.net/starcraft/Category:Strategies}}):
\begin{itemize}
  \item Protoss openings:
  \begin{itemize}
    \item \emph{two\_gates}: Build two \emph{Gateways} and keep training \emph{Zealots} (basic contact attack ground unit), this is the quickest way to apply pressure.
    \item \emph{fast\_dt}: Produce \emph{Dark Templars} (technologically advanced stealth ground unit) as soon as possible, sacrificing early game power for a technological advance, hard-countered by detectors technology.
    \item \emph{templar}: Train \emph{High Templars} (technologically advanced zone attack unit) as fast as possible, same as above, less deadly but is less easily countered.
    \item \emph{speedzeal}: Train \emph{Zealots} and research attack and speed upgrades as soon as possible, some early game power transitioning into late game tech.
    \item \emph{corsair}: Produce \emph{Corsairs} (air-air flying unit) as soon as possible and then transition into training \emph{Dark Templars} (safe from Zerg's flying detectors thanks to \emph{Corsairs}), \emph{Reavers} (ground artillery unit) or \emph{Dragoons}. Weak early game.
    \item \emph{nony}: Build three \emph{Gateways} and massive training of \emph{Dragoons}. Slower than \emph{two\_gates} but still some early game (ranged) power.
    \item \emph{reaver\_drop}: Train \emph{Reavers} as soon as possible to be able to do drops (air transport of artillery units).
  \end{itemize}
  \item Terran openings:
  \begin{itemize}
    \item \emph{bio}: Produce a large army of \emph{Marines} (basic ranged ground unit) and \emph{Medics} (can heal biological units). Quickest way to apply pressure.
    \item \emph{rax\_fe}: Take the closest ``natural expansion'' as soon as possible. This provides a big economic boost in the mid game by sacrificing some early game power.
    \item \emph{two\_facto}: Build two \emph{Factories} and keep producing \emph{Tanks} (ground artillery unit). Vulnerable while building up to it and then very powerful on ground.
    \item \emph{vultures}: Produce mainly \emph{Vultures} (fast ground ranged unit, excels against small units) and research mines. Quicker to reach (technologically) and build than tanks, can transition into tanks.
    \item \emph{air}: Produce \emph{Wraiths} (ranged flying units) as soon as possible for an air attack. Vulnerable to anti-air openings or quick rushes.
    \item \emph{drop}: Train \emph{Dropships} (flying transports) as soon as possible to be able to do (mostly tanks or marines) drops, leveraging efficient tactics.
  \end{itemize}
  \item Zerg openings:
  \begin{itemize}
    \item \emph{speedlings}: Train \emph{Zerlings} (basic cheap, fast, ground contact attack unit) and research speed upgrade as soon as possible. Quickest way to apply pressure.
    \item \emph{fast\_mutas}: Produce mainly \emph{Mutalisks} (ranged flying units). Vulnerable in the early game while gathering gas and researching the technology.
    \item \emph{mutas}: Expand two times for a stronger economy before massive training of \emph{Mutalisks}. Slower but more powerful build-up than above.
    \item \emph{lurkers}: Train \emph{Lurkers} (ground, stealth artillery unit) as soon as possible to benefit from their (advanced technology) zone attack and cloak ability. 
    \item \emph{hydras}: Massive production of \emph{Hydralisks} (ground ranged unit). Much quicker to reach technologically than \emph{Lurkers} and can transition into them.
  \end{itemize}
\end{itemize}

@@

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botopenings2011.png}
    \caption{Distribution of different openings performed by the different bots participating in the AIIDE 2011 competition. For each bot match-up, the colors show the proportion of times that the column bot used a particular opening against the row bot.}
    \label{fig:aiide2011-botopenings}
\end{figure*}

@@

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botopenings2012.png}
    \caption{Distribution of different openings performed by the different bots participating in the AIIDE 2012 competition. For each bot match-up, the colors show the proportion of times that the column bot used a particular opening against the row bot.}
    \label{fig:aiide2012-botopenings}
\end{figure*}

@@

\item {\bf Integration:}
%\begin{itemize}
%\item 
Multi-scale planning/reasoning: as described in this paper, all the bots developed for the StarCraft AI competitions decompose the problem of playing an RTS game into smaller subproblems, and then solutions for each of those subproblems are integrated in to a common architecture to play the game. However, the integration of each of the modules in a unified architecture is still an open problem. For example, how can decisions made at high-level modules be integrated with decisions made at lower-level modules? 
%\end{itemize}

\item {\bf Domain Knowledge: }
%\begin{itemize}
%\item 
We know how to incorporate some aspects of domain knowledge (e.g. build orders) into RTS game playing agents. But, in general, how to incorporate some forms of domain knowledge into algorithms for RTS games is still an open problem. For example, standard techniques to encode strategies for other forms of games, like Behavior Trees, are hard to deploy in RTS games. Is it possible to devise techniques that can automatically mine the existing collections of domain knowledge for an RTS game like StarCraft, and incorporate it into the bot? An initial exploration of this idea was carried out by Branavan et al. \cite{branavan2011learning}.
%\end{itemize}
\end{itemize}
